### 2011-04-04 Review/answers.tex
Made all Jos corrections.

\textbf{Answer:} The reviewer is right that accuracy of the method deteriorate with the high
viscosity ratio. However, the parameters we chose here as $\tau_{\mathrm{liq}}=3.5-7.5$ and with
grids in the vertical direction as $N_y=202$ is sufficient to suppress effects. We have added all
the necessary references in the text.

@@

\textbf{Answer:} We redrawed all the curves. For the figure 8 the data was extracted from the
figure and rescaled for the comparison reasons.

@@

\textbf{Answer:} The following section is added to the text 
\begin{quotation}
We chose to measure the film thickness in the middle of the bubble. The
classical Bretherton formulation is for the measurement of the film thickness at ``infinity''.
However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
it is enough for bubble to have a length of two-three diameters to apply analysis for very long
bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
channel heights from the front bubble tip. This distance was shown sufficient for the
film thickness to establish itself for the capillary number range $Ca<4$. We chose to measure
the film thickness in the middle of the bubble, which is located at least at the distance of $2.5-3$
channel heights from the front bubble tip.  Moreover, the film thickness examination, see Section
..., shows that for the capillary range of interest, i.e.
$0.08\leq Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
 percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones
($Ca \geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
height to the front tip. This certainly validates the point of choice to measure bubble thickness.
\end{quotation}

@@

\textbf{Question:}
\begin{quotation}
The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
Reynolds
number effects on the ﬁlm thickness for capillary numbers larger than 0.003” is in contradiction
with
the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
relatively
wide range of Reynolds numbers.”
\end{quotation}

\textbf{Answer:} The following paragraph is added:
\begin{quotation}
Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. In
the present simulations the largest Reynolds number we achieve is less than $20$ and it allows to
neglect the
inertia effects.}

\citet{kreutzer-pressure-drop} showed a deviation from the $Ca^{2/3}$ rule
{\color{red}\sout{and Reynolds number 
effects on the film thickness}} for capillary numbers larger than $0.003$.
\end{quotation}

@@

\textbf{Answer:} Thank you for pointing it out. We included the discussions of the Reynolds
effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
the capillary range regime and the grid depedency study.

@@

\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
mimics the force behavior.

@@

\textbf{Answer:} 
The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
current simulation equals to $7$, which is less than the number when the shape of the bubble is
affected \cite{zheng-large-ratio}. However, the point is that the equation of state in the lattice
Boltzmann framework is connected with density. Thus usually either pressure (i.e. density) boundary
conditions are prescribed or the body force. If they are prescribed simultaneously then
we can consider the influence of the Bond number on the simulations. Otherwise prescribing just the
body force is analogous to prescribing the pressure gradient $\frac{\mathrm{d}P}{\mathrm{d}x}$ in
the Navier-Stokes equation. Thus, it's not possbile the Bond number effect. It is because of the
coupled nature of the pressure and density in the LBM framework.

\textbf{Question:}
\begin{quotation}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estima
tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.
\end{quotation}

\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
about initialization and this was taken into the account by reducing repetitive text and
introducing step-by-step explations, see Section about initialization. We also added the
recommendations for controlling parameters in the same section.  The
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
keep suggestions for controlling and initialization of simulations as informative guides for the
LBM community.

\textbf{Question:}
\begin{quotation}
 Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
obtained
under the assumption of an axisymmetric Poiseuille ﬂow. However, among the ”Key words”, one can
read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
that are done in this work. The Authors should discuss this choice of estimating the pressure
gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
in Equation 5).
\end{quotation}

\textbf{Answer:} While the Hele-Shaw flow implies the flow with the infinitisemal distance between
plates, in our situation the size of the bubble can be compared with the channel height ($5$ times
larger) and that's why the assumption of the infinitisemal distance between plates is too strict.
Moreover, after recalculating results with the velocity of the interface the assumption of the
Poiseulle flow is valid and shows good accuracy.

\textbf{Question:}
\begin{quotation}
The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
journal of applied science. The Authors would improve signiﬁcantly the manuscript by deﬁning more
rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
A suggestion is to emphasize within an array the links between physical parameters and the model
eq
parameters. The variables i, fieq , gi should be deﬁned and the Equations for f and g in Equation
3 should be described (which equation corresponds to continuity equation...). Equation 2 could
perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
physical variable f of Equation 3. The notation i = 1 ÷ 8 is unusual, it would be clearer to use a
classical notation such as 1 ≤ i ≤ 8. The sound speed c2 should be deﬁned just after its
introduction in line 107.
\end{quotation}

\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
parameters with the lattice Botlzmann parameters. As well more explanations are put towards the
interface surface tension description and inclusion it to the model. The reference to the
continuity and the Navier-Stokes equations are added. The equation (2) is not rewritten in terms of
generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
governing equations. The notations are corrected.

@@

\textbf{Question:}
\begin{quotation}
The Authors should make a distinction between variables in lattice units and dimensional or non
dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y − 2$ where $N_y$ is
the
number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
is in
contradiction with the use of $H_{eff}$ in Equation 5. The Authors should deﬁne each variable
without
any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
improvements
would ﬁt well in the Subsection named ”The nondimensionalization and initialization procedure”
which does not in fact introduce much nondimensionalisation procedure.
\end{quotation}

\textbf{Answer:} We added a decription paragraph in the Subsection ``The nondimensionalization and
initialization procedure'' about the approach to units. In fact, all equations are the same in the
lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
preferred units as soon as capillary number is matched in the LBM and the physical world.

\textbf{Question:}
\begin{quotation}
Some notations are unclear. For instance, the ﬁlm thickness is denoted as h∞ in Equation 7 whereas
it is denoted δ in most of the paper. The notations w or δsim are also found before line 381 or in
Table 3 respectively. The thickness of the interface is denoted ξ in Figure 1 but the same variable
is
denoted 5ξ in Table 1. The speed of the bubble is denoted Ububble in some places and Ub in others.
The variable Uin Equation 5 does not seem to be deﬁned.
\end{quotation}

@@

\textbf{Question:}
\begin{quotation}
Three ﬁgures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
unusual
for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
which are not essential.
\end{quotation}

\textbf{Answer:} We eliminated necessary pictures and references to them.

@@

\textbf{Question:} 
\begin{quotation}
The plots of Figure 8 are of less quality than the other ﬁgures of the paper.
\end{quotation}

@@

\textbf{Question:}
\begin{quotation}
The work of Yang seems to be close from the present study since it involves the
solution of Bretherton
ﬂow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
differences between their work and this paper.
\end{quotation}

\textbf{Answer:} We outlined the differences in the introduction section.

@@

\textbf{Answer:} We added references to the introduction section.

@@

\textbf{Answer:} Thank you for pointing that. The classical Bretherton problem
$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
results are the same. It was found that Poiseuille profile assumption is improved significantly
after recalculation.

\textbf{Question:} 
\begin{quotation}
 If possible, the Authors should present results showing that steady state is reached. It would
also
be interesting if the Authors show some patterns of the ﬂow by plotting for instance the vector ﬁeld
of velocity around the bubbles.
\end{quotation}

\textbf{Answer:} We added an additional paragraph which shows the change of flow pattern for the
streamlines as it is indicated in literature.

\textbf{Question:}
\begin{quotation}
In the Abstract, the Authors should give the range of real capillary numbers they measured instead
of the targeted values (Calit ).
\end{quotation}

@@

\textbf{Answer:} We added additional information about results to the corresponding sections. The
pressure drop are not measured as they are influenced largely by Reynolds number which is beyond
the scope of this work.

@@

\textbf{Answer:} This is a part of discussion to have a remedy for the velocity. The corresponding
sentence is added.

@@

\textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
separate it from the numerical benchmark layout.

>>>>

\textbf{Answer:} The reviewer is right that accuracy of the method deteriorate with the high
viscosity ratio. However, the parameters we chose here as $\tau_{\mathrm{liq}}=3.5-7.5$ and with
grids in the vertical direction as $N_y=202$ is sufficient to suppress effects. We have added all
the necessary references in the text, lines $481-491$.

@@

\textbf{Answer:} We redrew all the figures. For the figure 8 the data was extracted from the
figure and rescaled for the comparison reasons.

@@

\textbf{Answer:} The explanations are added in lines 434-453.

@@

\textbf{Question:}
\begin{quotation}
The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
Reynolds
number effects on the film thickness for capillary numbers larger than 0.003” is in contradiction
with
the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
relatively
wide range of Reynolds numbers.”
\end{quotation}

\textbf{Answer:} We elaborate the Reynolds number influence in lines 34-43.

@@

\textbf{Answer:} Thank you for pointing this out. We included the discussions of the Reynolds
effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
the capillary range regime and the grid depedency study.

@@

\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
how it
mimics the force behavior.

@@

\textbf{Answer:} 
The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
current simulation equals to $7$, which is less than the number for which the shape of the bubble is
affected according to \cite{zheng-large-ratio}.

\textbf{Question:}
\begin{quotation}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estimations of the body force to apply and of the film thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.
\end{quotation}

\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
about initialization and this was taken into account by reducing repetitive text and
introducing step-by-step explations, see Section about initialization (lines 259-305). We also added
the
recommendations for controlling parameters in the same section.  The
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann method are refered to the non-dimensionalization procedure. Therefore, we
decided to
keep suggestions for controlling and initialization of simulations as informative guidelines for the
LBM community.

\textbf{Question:}
\begin{quotation}
 Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
obtained
under the assumption of an axisymmetric Poiseuille flow. However, among the ”Key words”, one can
read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
that are done in this work. The Authors should discuss this choice of estimating the pressure
gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
in Equation 5).
\end{quotation}

\textbf{Answer:} The factor $1/8$ is taken for the maximum velocity for the Poiseille flow between
plates. Our assumption that the bubble travels with the constant velocity which equals to the
maximum velocity of the Poiseuille profile between plates. Factor $1/12$ comes from the averaging
of the flow rate. However, bubble moves faster than the liquid, i.e. the average liquid velocity.
That's why the factor $1/8$ was chosen. The simulations show that factor $1/8$ works reasonably
well.

\textbf{Question:}
\begin{quotation}
The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
journal of applied science. The Authors would improve signiﬁcantly the manuscript by defining more
rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
A suggestion is to emphasize within an array the links between physical parameters and the model
parameters. The variables $i$, $f_i^{eq}$ , $g_i$ should be deﬁned and the Equations for $f$ and $g$
in Equation
3 should be described (which equation corresponds to continuity equation...). Equation 2 could
perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
physical variable f of Equation 3. The notation $i = 1 \div 8$ is unusual, it would be clearer to
use a
classical notation such as $1 \leq i \leq 8$. The sound speed $c_s^2$ should be deﬁned just after
its
introduction in line 107.
\end{quotation}

\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
parameters with the lattice Botlzmann parameters (lines $149-160$). As well more explanations are
put towards the
interface surface tension description (lines $136-139$). The reference to the
continuity and the Navier-Stokes equations are added (lines $140-143$). The equation (2) is not
rewritten in terms of
generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
governing equations. The notations have been corrected.

@@

\textbf{Question:}
\begin{quotation}
The Authors should make a distinction between variables in lattice units and dimensional or non
dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y - 2$ where $N_y$ is
the
number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
is in
contradiction with the use of $H_{eff}$ in Equation 5. The Authors should define each variable
without
any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
improvements
would fit well in the Subsection named ”The nondimensionalization and initialization procedure”
which does not in fact introduce much nondimensionalisation procedure.
\end{quotation}

\textbf{Answer:} We added a decription paragraph about the approach to units (lines 149-160). In
fact, all equations are the same in the
lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
preferred units as soon as the capillary number is the same in the LBM and the physical world.

\textbf{Question:}
\begin{quotation}
Some notations are unclear. For instance, the film thickness is denoted as $h_{\infty}$ in Equation
7 whereas
it is denoted $\delta$ in most of the paper. The notations w or $\delta_{sim}$ are also found before
line 381 or in
Table 3 respectively. The thickness of the interface is denoted $\xi$ in Figure 1 but the same
variable
is
denoted $5\xi$ in Table 1. The speed of the bubble is denoted $U_{\mathrm{bubble}}$ in some places
and $U_b$ in others.
The variable $Uin$ Equation 5 does not seem to be deﬁned.
\end{quotation}

@@

\textbf{Question:}
\begin{quotation}
Three figures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
unusual
for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
which are not essential.
\end{quotation}

\textbf{Answer:} We eliminated unnecessary pictures and references to them.

@@

\textbf{Question:} 
\begin{quotation}
The plots of Figure 8 are of less quality than the other figures of the paper.
\end{quotation}

@@

\textbf{Question:}
\begin{quotation}
The work of Yang seems to be close from the present study since it involves the
solution of Bretherton
flow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
differences between their work and this paper.
\end{quotation}

\textbf{Answer:} We outlined the differences in the introduction section (lines $92-100$).

@@

\textbf{Answer:} We added references to the introduction section [17-19].

@@

\textbf{Answer:} Thank you for pointing that out. In the classical Bretherton problem
$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
results are the same.

\textbf{Question:} 
\begin{quotation}
 If possible, the Authors should present results showing that steady state is reached. It would
also
be interesting if the Authors show some patterns of the flow by plotting for instance the vector
ﬁeld
of velocity around the bubbles.
\end{quotation}

\textbf{Answer:} We added a small steady-state study (lines $334-341$). We added an additional
paragraph which shows the change of flow pattern for the
streamlines as it is indicated in literature (lines $463-471$, Figure 8).

\textbf{Question:}
\begin{quotation}
In the Abstract, the Authors should give the range of real capillary numbers they measured instead
of the targeted values ($Ca_{lit}$).
\end{quotation}

@@

\textbf{Answer:} We added additional information about results to the corresponding sections. The
pressure drop is not measured as they are influenced largely by Reynolds number according to
\citet{heil-bretherton}.

@@

\textbf{Answer:} This is a part of discussion. The corresponding
sentence is added (lines $459-461$).

@@

\textbf{Answer:} The corresponding figure for the classical Bretherton problem is added. We
separated it from the numerical benchmark layout.
